Technical Interview

### 2017-09-06 tex/paper/paper.tex
Fix paper typo
version 1.1

\newlength\drop
\makeatletter
\newcommand*\titleM{\begingroup%
\setlength\drop{0.08\textheight}
\centering
\vspace*{\drop}
{\Huge\bfseries Technical Interview}\\[\baselineskip]
%{\scshape MAN-AHL}\\[\baselineskip]
\vfill
\includegraphics[width=0.6\textwidth]{img/logo.png}\par\vspace{1cm}
\vfill
{\Huge\scshape Fadil Mokhchane \\ 
\small \href{https://www.linkedin.com/in/fadil/}{CV on LinkedIn} }\par
\vspace{1cm}
\includegraphics[width=0.3\textwidth]{img/harrington.jpg}\par
\vfill
\vspace*{2\drop}
{\scshape September 3, 2017}\par
{\scshape \small Version 1.0}\par
\endgroup}
\makeatother
%% END Memoir customization

@@

\newpage
\tableofcontents* % the asterisk means that the contents itself isn't put into the ToC
%\clearpage %or \cleardoublepage
\phantomsection
\listoffigures
\listoftables
\vfill
\newpage

@@

\subsection{Design}
We target the following archictecture, c.f. \autoref{fig:architecture1}.
\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.8\textwidth]{img/architecture1.png}
\caption{Algorithm App Architecture}
\label{fig:architecture1}
\end{center}
\end{figure}
The various modules implement all the functionalities we need.
\begin{description}
	\item[Analytics] The mathematical functions required,

@@

	This module contains an abstraction to a uniform pseudo
	random number generator with values bounded 
	within $ \left[ 0, 1 \right]$.
	It can compute a random number or a series of random 
	number and the statistics related to the distribution of 
	the random number generated.
	\item[Weighted Engine] 	The weighted pseudo random 
	numbers probability engine,

	This module includes an abstraction to the weighted 
	pseudo random numbers generators.
	It allows computation and statistics on these numbers.
	It includes the requested nextNum function.
	\item[CLI] 		The command line input layer,

@@

\subsection{Uniform Engine Module}

The \emph{UniformRNG} ADT abstracts a choice of uniform pseudo random 
number generator, Mersenne Twister and Ecuyer. We could seemlessly
add new generator, if needs be. We would need to extend the standard
Haskell \emph{RandomGen} class. 
They can generate un-bounded pseudo random numbers.

@@

The \emph{StatUniEngine} ADT abstracts the computation of the statistics
related to the bounded uniform random numbers generated. 
In order to be efficient, we compute only the distribution while we 
are creating pseudo random numbers. 
There is a post-processing step at the end of the simulation, which 
computes the remaining relevant statistics.
It uses the \emph{State} monad to pass on the state of the 
uniform random number generator and the cumulative statistics.
This engine instanciates the \emph{StatEngine} class, which role is
to compute the statistics of a series of random numbers.

\subsection{Weighted Engine Module}

The \emph{ProbaWPEngine} ADT allows to computed weighted pseudo
random numbers. It order to achieve this, it need to build and keep
thoughout the computations the \emph{EngineParams} ADT. 
This latter encapsulates the \emph{PDF}, \emph{CDF} and \emph{InvCDF}
ADTs. They do not change thoughout the computations and they are kept 
in memory for efficiency. We pass them thought the computations
using the \emph{Reader} monad. We also use the \emph{State} monad
to pass on the state of the bounded uniform engine.
It also instanciates the \emph{ProbaEngine} class.

The \emph{StatWPEngine} ADT compute the statistic for a weighted
probabilities run. It also use a monad composed of a \emph{Reader}
and \emph{State} monads. A post processing step computes
the remaining metrics. It also instanciates the \emph{StatEngine} class.

@@

%-------------------------------------------------------------------------------
\section{Performance Analysis}

@@

\subsection{Uniform Engine Performance}
\autoref{tab:uni.perf} presents the result of many runs with various $n$ and
the heap that was allocated during the run and the time taken.
\begin{table}[htbp!]
  \centering \tiny
    \begin{tabular}{r|rrrrrr}
    \multicolumn{1}{r}{} & \multicolumn{2}{c}{1000000} & \multicolumn{2}{c}{100000} & \multicolumn{2}{c}{10000} \\
    \multicolumn{1}{r}{} & Mersenne & Ecuyer & Mersenne & Ecuyer & Mersenne & Ecuyer \\
\cmidrule{2-7}    Heap  & 2,724,561,112 & 3,360,007,920 & 274,635,272 & 338,190,648 & 29,602,472 & 35,953,944 \\
    Time  & 3.576 & 3.707 & 0.347 & 0.373 & 0.041 & 0.041 \\
    \end{tabular}%
  \caption{Uniform Engine Performance}
  \label{tab:uni.perf}%
\end{table}%
Most of the heap allocation and the time is consumed by the underlying 
random number generator, the various engines adds only minor overhead, 
as shown on the profiling results, c.f. \autoref{fig:perf1} and \autoref{fig:perf2}.
We can see that the time and space are linear with respect to $n$. 
\begin{figure}[h!]
\begin{center}
\includegraphics[width=1\textwidth]{img/prof1.png}
\caption{Performance Profiling 1}
\label{fig:perf1}
\end{center}
\end{figure}
\begin{figure}[h!]
\begin{center}
\includegraphics[width=1\textwidth]{img/prof2.png}
\caption{Performance Profiling 2}
\label{fig:perf2}
\end{center}
\end{figure}
%\begin{figure}[h!]
%\begin{center}
%\includegraphics[width=1\textwidth]{img/prof3.png}
%\caption{Performance Profiling 3}
%\label{fig:perf3}
%\end{center}
%\end{figure}
We can conclude that the complexity of the inner random number generators
are $O(1)$ in time and $O(n)$ in space. Indeed the function \emph{allStats} for the
\emph{StatUniEngine} does not store the number generated, 
just the distribution of these.

@@

\autoref{tab:w.perf} presents similar results for the \emph{StatWPEngine}, 
i.e. the statistic weighted probability engine.
\begin{table}[htbp!]
  \centering \tiny
    \begin{tabular}{r|rrrrrr}
    \multicolumn{1}{r}{} & \multicolumn{2}{c}{1000000} & \multicolumn{2}{c}{100000} & \multicolumn{2}{c}{10000} \\
    \multicolumn{1}{r}{} & Mersenne & Ecuyer & Mersenne & Ecuyer & Mersenne & Ecuyer \\
\cmidrule{2-7}    Heap  & 2,777,630,784 & 3,419,213,784 & 281,175,032 & 344,423,280 & 30,695,560 & 37,067,344 \\
    Time  & 3.612 & 3.81  & 0.357 & 0.379 & 0.038 & 0.043 \\
    \end{tabular}%
  \caption{Weighted Engine Performance}
  \label{tab:w.perf}%
\end{table}%
The results shows that there is little overhead for the weighted probabilities
engines. The results are very similar to \autoref{tab:uni.perf}.
This conforts us in the computation of our complexities.

%-------------------------------------------------------------------------------
\section{Possible Extensions}
Here is a list of possible extension we could add.
\begin{description}
	\item [Analytics]
	Further abstract the PDF, CDF, InvCDF to handle all kind of keys.

	Review the design of the statistics.
	\item [Testing]
	We could run the tests on different environments to ensure the portability of 
	the code.
	\item [UI] 
	A nice extension would be to add a Graphic User Interface,
	using the library 
	\href{https://hackage.haskell.org/package/threepenny-gui}{threepenny-gui}
	to create a portable Web application.

@@

\end{document}

>>>>

\newlength\drop
\makeatletter
\newcommand*\titleM{\begingroup%
\setlength\drop{0.08\textheight}
\centering
\vspace*{\drop}
{\Huge\bfseries Technical Interview}\\[\baselineskip]
%{\scshape MAN-AHL}\\[\baselineskip]
\vfill
\includegraphics[width=0.6\textwidth]{img/logo.png}\par\vspace{1cm}
\vfill
{\Huge\scshape Fadil Mokhchane \\ 
\small \href{https://www.linkedin.com/in/fadil/}{CV on LinkedIn} }\par
\vspace{1cm}
\includegraphics[width=0.3\textwidth]{img/harrington.jpg}\par
\vfill
\vspace*{2\drop}
{\scshape September 6, 2017}\par
{\scshape \small Version 1.1}\par
\endgroup}
\makeatother
%% END Memoir customization

@@

\newpage
\tableofcontents* % the asterisk means that the contents itself isn't put into the ToC
%\clearpage %or \cleardoublepage
\phantomsection
\vfill

\begin{table}[h!]
  \centering
  \caption{Table of Changes}
%  \label{tab:table1}
  \begin{tabular}{lll p{6.cm}}
    \toprule
    V. & Author & Date & Change\\
    \midrule
    1.0 & Fadil M. & 3/09/2017 & 
    	Initial version. \\
    1.1 & Fadil M. & 6/09/2017 & 
    	Simplified and generalised the design of the code. 
    	C.f. \autoref{apx:code}

    	- Added the Statistics engine module in \autoref{ssec:design}

    	- Updated implementation notes: \autoref{ssec:uniform.impl}
    	 and \autoref{ssec:weighted.impl}, 
    	added \autoref{ssec:stats.impl} to reflect the new design.

    	- Updated the performances, \autoref{sec:perf}, to be clearer.

    	- Updated the extensions, \autoref{sec:ext}.

    	- Added \autoref{apx:code} with the code to facilitate sharing.
    	\\
    \bottomrule
  \end{tabular}
\end{table}

\newpage
\listoffigures
\listoftables
\vfill
\newpage

@@

\subsection{Design}
\label{ssec:design}
We target the following archictecture, c.f. \autoref{fig:architecture1}.
\begin{figure}[h!]
\begin{center}
\includegraphics[width=0.9\textwidth]{img/architecture1.png}
\caption{Algorithm App Architecture}
\label{fig:architecture1}
\end{center}
\end{figure}
The various modules implement all the functionalities we need.
\begin{description}
	\item[Analytics] The mathematical functions required,

@@

	This module contains an abstraction to a uniform pseudo
	random number generator with values bounded 
	within $ \left[ 0, 1 \right]$.
	It can compute a random number or a series of random 
	number.

	\item[Weighted Engine] 	The weighted pseudo random 
	numbers probability engine,

	This module includes an abstraction to the weighted 
	pseudo random numbers generators.
	It allows computation and statistics on these numbers.
	It includes the requested nextNum function.
	\item[Statistics Engine] The various statistical metrics
	introduced in \autoref{sec:quality}.

	It computes the statistics metrics for a series of 
	random numbers produced by the weighted or
	uniform engine.
	\item[CLI] 		The command line input layer,

@@

\subsection{Uniform Engine Module}
\label{ssec:uniform.impl}
The \emph{UniformRNG} ADT abstracts a choice of uniform pseudo random 
number generator, Mersenne Twister and Ecuyer. We could seemlessly
add new generator, if needs be. We would need to extend the standard
Haskell \emph{RandomGen} class. 
They can generate un-bounded pseudo random numbers.

@@

\subsection{Weighted Engine Module}
\label{ssec:weighted.impl}
The \emph{ProbaWPEngine} ADT allows to computed weighted pseudo
random numbers. It order to achieve this, it need to build and keep
thoughout the computations the \emph{EngineParams} ADT. 
This latter encapsulates the \emph{PDF}, \emph{CDF} and \emph{InvCDF}
ADTs. They do not change thoughout the computations and they are kept 
in memory for efficiency. We pass them thought the computations
using the \emph{Reader} monad. We also use the \emph{State} monad
to pass on the state of the bounded uniform engine.
It also instanciates the \emph{ProbaEngine} class.

\subsection{Statistics Engine Module}
\label{ssec:stats.impl}
This \emph{StatEngine} type role is
to compute the statistics of a series of random numbers generated
by either the \emph{ProbaUniEngine} or the \emph{ProbaWPEngine}.

In order to be efficient, we compute only the distribution while we 
are creating pseudo random numbers. 
There is a post-processing step at the end of the simulation, which 
computes the remaining relevant statistics.
It uses the \emph{State} monad to pass on the state of the 
uniform random number generator and the cumulative statistics.
It also use a monad composed of a \emph{Reader} which
pass one the engine used and the engine parameters.

@@

%-------------------------------------------------------------------------------
\section{Performance Analysis}
\label{sec:perf}

@@

\subsection{Uniform Engine Performance}
\autoref{tab:uni.perf} and \autoref{fig:perf.uni} 
 present the result of many runs with various $n$ and
the heap (in bytes) that was allocated during the run and the time taken.
\begin{figure}[h!]
\centering
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=\textwidth]{img/perf-heap-uni.png}
        \caption{Heap (bytes)}
        \label{fig:perf.heap.uni}
    \end{subfigure}
    ~ 
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=\textwidth]{img/perf-time-uni.png}
        \caption{Time (s)}
        \label{fig:perf.time.uni}
    \end{subfigure}
\caption{Performance Uniform Engine}
\label{fig:perf.uni}
\end{figure}
\begin{table}[htbp!]
  \centering \tiny
    \begin{tabular}{r|rrrrrr}
    \multicolumn{1}{r}{} & \multicolumn{2}{c}{1000000} & \multicolumn{2}{c}{100000} & \multicolumn{2}{c}{10000} \\
    \multicolumn{1}{r}{} & Mersenne & Ecuyer & Mersenne & Ecuyer & Mersenne & Ecuyer \\
\cmidrule{2-7}    Heap  & 2,724,561,112 & 3,360,007,920 & 274,635,272 & 338,190,648 & 29,602,472 & 35,953,944 \\
    Time  & 3.576 & 3.707 & 0.347 & 0.373 & 0.041 & 0.041 \\
    \end{tabular}%
  \caption{Uniform Engine Performance}
  \label{tab:uni.perf}%
\end{table}%

Most of the heap allocation and the time is consumed by the underlying 
random number generator, the various engines adds only minor overhead, 
as shown on the profiling results, c.f. the file \emph{Run.prof} in the code.
We can see that the time and space are linear with respect to $n$.

We can conclude that the complexity of the inner random number generators
are $O(1)$ in time and $O(n)$ in space. Indeed the function \emph{allStats} for the
\emph{StatUniEngine} does not store the number generated, 
just the distribution of these.

@@

\autoref{tab:w.perf} presents similar results for the \emph{StatWPEngine}, 
i.e. the statistic weighted probability engine.
\begin{figure}[h!]
\centering
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=\textwidth]{img/perf-heap-w.png}
        \caption{Heap (bytes)}
    \end{subfigure}
    ~ 
    \begin{subfigure}[b]{0.47\textwidth}
        \includegraphics[width=\textwidth]{img/perf-time-w.png}
        \caption{Time (s)}
    \end{subfigure}
\caption{Performance Weighted Engine}
\label{fig:perf.w}
\end{figure}
\begin{table}[htbp!]
  \centering \tiny
    \begin{tabular}{r|rrrrrr}
    \multicolumn{1}{r}{} & \multicolumn{2}{c}{1000000} & \multicolumn{2}{c}{100000} & \multicolumn{2}{c}{10000} \\
    \multicolumn{1}{r}{} & Mersenne & Ecuyer & Mersenne & Ecuyer & Mersenne & Ecuyer \\
\cmidrule{2-7}    Heap  & 2,777,630,784 & 3,419,213,784 & 281,175,032 & 344,423,280 & 30,695,560 & 37,067,344 \\
    Time  & 3.612 & 3.81  & 0.357 & 0.379 & 0.038 & 0.043 \\
    \end{tabular}%
  \caption{Weighted Engine Performance}
  \label{tab:w.perf}%
\end{table}%
The results shows that there is little overhead for the weighted probabilities
engines. The results are very similar to \autoref{tab:uni.perf}.
This conforts us in the computation of our complexities.

%-------------------------------------------------------------------------------
\section{Possible Extensions}
\label{sec:ext}
Here is a list of possible extension we could add.
\begin{description}
	\item [Testing]
	We could run the tests on different environments to ensure the portability of 
	the code.
	\item [UI] 
	A nice extension would be to add a Graphic User Interface,
	using the library 
	\href{https://hackage.haskell.org/package/threepenny-gui}{threepenny-gui}
	to create a portable Web application.

@@

%-------------------------------------------------------------------------------
\newpage
\appendix
\addcontentsline{toc}{chapter}{Appendices}

\chapter{Code} \label{apx:code}
We provide below the code. We use Vim with the extension 
\href{https://github.com/begriffs/haskell-vim-now}{haskell-vim-now}
\end{document}
